---
title: "DSC520_Week3.1_Guruprasad_Velikadu_Assignment03"
author: "Guruprasad Velikadu Krishnamoorthy"
date: "2022-12-18"
output: 
  pdf_document: 
    toc: yes
    number_sections: yes
    highlight: espresso
    
---

# Assignment

## Load the ggplot2 package
```{r}
library(ggplot2)
theme_set(theme_minimal())
library(formatR)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```

## Set the working directory to the root of your DSC 520 directory
```{r}
knitr::opts_knit$set(root.dir ="C:/Users/Gurup/GURU/Learning/Masters/Term_2/DSC520_T302_Statistics_for_Data_Science/Week_1/forked_example/dsc520_Week_1_Assignment_forkedExample")
```



## Load the `data/r4ds/heights.csv` to
```{r}
heights_df <- read.csv("data/r4ds/heights.csv")
```



## https://ggplot2.tidyverse.org/reference/geom_point.html
### Using `geom_point()` create three scatterplots for
### `height` vs. `earn`
```{r}
ggplot(data=heights_df, aes(x=height, y=earn)) + geom_point(size=3)+theme_bw()+labs(title= "height vs earn")
```



## `age` vs. `earn`
```{r}
ggplot(data=heights_df, aes(x=age, y=earn)) + geom_point(size=3)+theme_bw()+labs(title= "age vs earn")
```


## `ed` vs. `earn`
```{r}
ggplot(data=heights_df, aes(x=ed, y=earn)) + geom_point(size=3)+theme_bw()+labs(title= "ed vs earn")
```



## Re-create the three scatterplots and add a regression trend line using
### the `geom_smooth()` function
### `height` vs. `earn`
```{r}
ggplot(data=heights_df, aes(x=height, y=earn)) + geom_point(size=3) +geom_smooth()+theme_bw()+labs(title= "height vs earn")
```

## `age` vs. `earn`
```{r}
ggplot(data=heights_df, aes(x=age, y=earn)) + geom_point(size=3) + geom_smooth()+theme_bw()+labs(title= "age vs earn")
```



## `ed` vs. `earn`
```{r}
ggplot(data=heights_df, aes(x=ed, y=earn)) + geom_point(size=3) + geom_smooth(size=2)+theme_bw()+labs(title= "ed vs earn")
```



## Create a scatterplot of `height`` vs. `earn`.  Use `sex` as the `col` (color) attribute
```{r}
ggplot(data=heights_df, aes(x=height, y=earn, col=sex)) + geom_point(size=3)+theme_bw()+labs(title= "height vs earn")
```



## Using `ggtitle()`, `xlab()`, and `ylab()` to add a title, x label, and y label to the previous plot
### Title: Height vs. Earnings
### X label: Height (Inches)
### Y Label: Earnings (Dollars)
```{r}
ggplot(data=heights_df, aes(x=height, y=earn, col=sex)) + geom_point(size=3) + ggtitle("Height vs. Earnings") + xlab("Height (Inches)") + ylab("Earnings (Dollars)")+theme_bw()
```



## https://ggplot2.tidyverse.org/reference/geom_histogram.html
### Create a histogram of the `earn` variable using `geom_histogram()`
```{r}
ggplot(data=heights_df, aes(x=earn)) + geom_histogram()+theme_bw()+labs(title= "Earnings")
```



## Create a histogram of the `earn` variable using `geom_histogram()`
### Use 10 bins
```{r}
ggplot(data=heights_df, aes(x=earn)) + geom_histogram(bins = 10)+theme_bw()+labs(title= "Earnings")
```



## https://ggplot2.tidyverse.org/reference/geom_density.html
### Create a kernel density plot of `earn` using `geom_density()`
```{r}
ggplot(data=heights_df, aes(x=earn)) +  geom_density(size=1)+theme_bw()+labs(title= "Earnings")
```

# Session Info
```{r}
sessionInfo()
```



